\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}

### 2021-10-05 lin-vs-cen.tex
Almost finished (modulo boundary issue) under <=r assumption

For a graph $G$ and a vertex $v\in V(G)$, $\deg_G(v):=|\{vw\in E(G)\}|$ denotes the degree of $v$ in $G$.  For a any $S\subseteq V(G)$, $\deg_G(S):=\sum_{v\in S}\deg_G(v)$ is the total degree of $S$.\todo{Surely this next lemma is well-known?}

\begin{lem}\label{hall_vees}
  Let $G$ be a bipartite graph with vertex parts $A$ and $B$ and having the property that, for any $A_0\subseteq A$, $|N_G(W)|\ge 2|A_0|$.  Then $G$ contains a subgraph $\Lambda$ in which each vertex of $A$ has degree exactly two and each vertex of $B$ has degree at most one.
\end{lem}

@@

  Let $A_0\subseteq A$ and $B_0\subseteq B$ be the subsets of $A$ and $B$, respectively, that can be reached from $v_0$ by $\Lambda$-alternating paths.
  Observe that, for any $x\in B_0$,  there is exactly one edge $vx$ of $\Lambda$ incident on $x$ and the other endpoint $v$ of this edge is in $A_0$.  Let $a$ be the number of edges of $\Lambda$ having one endpoint $A_0$ and one endpoint in $B_0$, so that $a=\deg_\Lambda(B_0)=|B_0|$.  Let $b$ be the number of edges of $\Lambda$ having one endpoint in $A_0$ and one endpoint in $B\setminus B_0$, so that $\deg_\Lambda(A_0)=a+b = |B_0|+b$.    Since $\deg_{\Lambda}(v) \le 2$ for each $v\in A_0$ and $\deg_\Lambda(v_0)<2$, $\deg_\Lambda(A_0)<2|A_0|$.  This yields the desired contradiction, since $|N_\Lambda(A_0)| \le \deg_\Lambda(A_0) = |B_0|+b = \deg_\Lambda(A_0)< 2|A_0|$.
\end{proof}

@@

A \emph{subdivision} of a graph $G$ is any graph $G'$ that can be obtained from $G$ by replacing each edge $vw$ of $G$ with a path $v,s_1,s_2,\ldots,s_\ell,w$ in which each of the (newly introduced) internal vertices $s_1,\ldots,s_\ell$ have degree exactly $2$.  Each vertex in $V(G')\setminus V(G)$ is called a \emph{subdivision vertex}.  A $\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has exactly $\ell$ internal vertices.  For any graph $G$, we use $G^{\onesub}$ to denote the $1$-subdivision of $G$.  Note that there is a natural bijection between the vertex set $V(G^{\onesub})$ and the vertex/edge set $V(G)\cup E(G)$.

@@

\begin{proof}
  Let $P:=\{v_1,\ldots,v_t\}$ where $p_i^{\leftup}:=(x_i,y_i)$.  Let $X:=\bigcup_{i=1}^t \{x_i\bmod 10,(x_{i+1})\bmod 10\}$ and let $Y:=\bigcup_{i=1}^t\{y_i\bmod 10,(y_i+1)\bmod 10\}$.  Then $|X|,|Y|\le 2t\le 6$.  In particular, there exists some $x\in\{0,\ldots,6\}\setminus X$ and some $y\in\{0,\ldots,6\}\setminus Y$. For each $i,j\in\Z$, let $c_{i,j}:=(7i+x+3,7j+y+3)$ and observe that the ball $B_{3}(c_{i,j})$ does not have any point of $P$ on its boundary.  Furthermore, the set of balls $S:=\{B_{3}(c_{i,j}):i,j\in\Z\}$ is a partition of $V(G^{\onesub}_{k})$.  Therefore $S$ is a set of pairwise-disjoint balls, each of radius $3$, such that each vertex in $P$ is contained in the interior of some ball in $S$.
\end{proof}

\begin{lem}
  Let $P$ be a subset of $V(G^{\onesub}_k)$ such that no ball of radius $9$ contains more than $3$ vertices in $P$.  Then there exists a set $S$ of disjoint balls, each of radius $3$ and such that each point of $P$ is in the interior of some ball in $S$.
\end{lem}

\begin{proof}
  Define a graph $H$ with vertex set $V(H):=P$ and that contains an edge $pq$ if and only if $\|pq\|_\infty \le 12$. Observe that, if some connected subgraph of $H$ contains $4$ vertices, then these $4$ vertices are contained in a ball of radius $9$.  Therefore each component of $H$ contains at most $3$ vertices.

@@

\begin{lem}\label{one_path_enchilada}
  Let $k$ be a positive integer, let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{k})\setminus\partial G^{\onesub}_k$ that have the following properties:
  \begin{compactenum}[(1)]\setcounter{enumi}{-1}
    \item Any ball of radius 9 contains at most $3$ elements of $I\cup L\cup Q$.
    \item Each vertex in $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
    \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
    \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
  \end{compactenum}
  Then $G^{\onesub}_k$ contains a path $P$ that
  \begin{compactenum}[(a)]
    \item contains all the vertices in $I\cup L\cup Q$;
    \item goes straight through each vertex in $I$; and
    \item turns at each vertex in $L$.
  \end{compactenum}
\end{lem}

\begin{proof}
  Start with the snakelike path $P_0$ that visits every $7$th row of $G_k$. Cover the elements of $I\cup L\cup Q$ with disjoint balls of side length $3$.  Each such square is intersected by exactly one row of the snakelike path.  Use \cref{seven_by_seven} to locally deform $P_0$ to pick up the objects in any square that it intersects.
\end{proof}

@@

  We now argue that $\vor'$ satisfies (\ref{three_in_ball}).  Let $v_1,\ldots,v_4\in Q$ and $1\le i_1<\cdots<i_4\le|P|$ be such that $\vor'(v_j)=p_{i_j}$ for each $j\in\{1,\ldots,4\}$.  Suppose, for the sake of contradiction, that $i_1,\ldots,i_4$ are all distinct and that $v_1,\ldots,v_4$ are contained in some ball $B_{\beta/12}(c)$.  By the triangle inequality $\|v_iv_j\|_\infty \le \|v_ic\|_\infty+\|cv_j\|_\infty \le \beta/6$ for each  $i,j\in\{1,\ldots,4\}$.
  Since $v_1\in Q$, $d(v)\le\beta$ so
  \[ 
    \|p_{i_1}v_1\|_\infty \le d(v_1) + 3\beta/4 \le \beta + 3\beta/4 \enspace .
  \]
  Since $\vor'(v_2)=p_{i_2}\neq p_{i_1}$,
  \[ 
    \|p_{i_2}v_2\|_\infty
     < \|p_{i_1}v_2\|_\infty - 3\beta/4
     \le \|p_{i_1}v_1\|_\infty + \beta/6 - 3\beta/4
     \le \beta + \beta/5 \enspace .
  \]
  Since $\vor'(v_3)=p_{i_3}\neq p_{i_2}$,
  \[
     \|p_{i_2}v_3\|_\infty
     < \|p_{i_2}v_3\|_\infty - 3\beta/4
     \le \|p_{i_2}v_2\|_\infty + \beta/6 - 3\beta/4
     < \beta + 2\beta/6 - 3\beta/4 \enspace .
  \]
  Finally, since $\vor'(v_4)=p_{i_4}\neq p_{i_3}$,
  \[
    \|p_{i_4}v_4\|_\infty
    < \|p_{i_3}v_4\|_\infty - 3\beta/4
    \le \|p_{i_3}v_3\|_\infty + \beta/5 - 3\beta/4
    < \beta + 3\beta/6 - 3\beta/2 = 0 \enspace ,
  \]
  which is clearly a contradiction.  This establishes that $\vor'$ satisfies Condition~(\ref{three_in_ball}).\todo{Optimize and cleanup. Finish proving (iii)}\todo{Why not use $\beta$ (or $100\beta$) instead of $3\beta/4$?}
\end{proof}

@@

\begin{proof}
  (This proof only uses the fact that $G_{a\times b})$ has no vertex of degree greater than $4$.)  Since $G$ contains a $G_{a\times b}$ minor then, by definition, $G$ contains a subgraph $H$ and a partition $\mathcal{B}:=(B_x:x\in V(G_{a\times b}))\}$ of $V(H)$ such that
  \begin{inparaenum}[(i)]
    \item $B_x$ is non-empty and $H[B_x]$ is connected for each $x\in V(G_{a\times b})$;
    \item $H/\mathcal{B}$ is isomorphic to $G_{a\times b}$ using the isomorphism given by $B_x\mapsto x$.
  \end{inparaenum}
  Let $(H,\mathcal{B})$ be chosen so that $H$ the minimum possible number of edges, over all choices $H$ and $\mathcal{B}$ that satisfy (i) and (ii).

  An edge of $vw$ of $H$ is a \emph{border edge} if $v\in B_x$ and $w\in B_y$ for distinct $x,y\in V(G_{a\times b})$.  For each $x\in V(G_{a\times b})$, let $\partial B_x$ denote the set of border edges incident on one vertex of $B_x$.  Since $H$ is edge-minimal, $|\partial B_x|=\deg_{G_{a\times b}}(x)\le 4$, for each $x\in V(G_{a\times b})$.  For each $x\in V(G_{a\times b})$, let $T_x\subseteq H$ be a minimal tree contained in $H[B_x]\cup \partial B_x$ that contains all the border edges in $\partial B_x$.   We define the partition $\mathcal{P}:=(P_x:x\in V(G_{a\times b}))$ as follows:  
\begin{compactenum}
  \item If $T_x$ has a node $v$ of degree $4$, then $P_x:=\{v\}$.  
  \item If $T_x$ has two nodes $v$ and $w$ of degree $3$ then $P_x$ contains all vertices on the path in $T_x$ from $v$ to $w$.
  \item If $T_x$ has exactly one node $v$ of degree $3$ then $P_{x}:=\{v\}$
  \item Otherwise $T_x$ is a path and $P_x$ contains exactly one node in $V(T_x)$.
\end{compactenum}
It is straightforward to verify that, with this assignment $\mathcal{P}:=(P_x:x\in V(G_{a\times b}))$, the quotient graph $H/\mathcal{P}$ is a subdivision of $G_{a\times b}$, so that $H$ is indeed an $a\times b$ pseudogrid.
\end{proof}

The second case (ii) in the proof of \cref{pseudogrid_minor} occurs when $\deg_{G_{a\times b}}(x)=4$, so $x=(i,j)$ for some $i\in\{2,\ldots,a-1\}$ and $j\in\{2,\ldots,b-1\}$.  This case is especially important in our proof because it corresponds to a case in which some of the vertices of $P_x$ are shared by the $i$th column and the $j$th row of the pseudogrid. However, a path in the pseudogrid that uses vertices of $B_x$ may or may not contain all the vertices of $P_x$ depending on whether or not the corresponding path in $G_{a\times b}$ bends at $x$.  This motivates the following definitions (see \cref{straight_or_bent}):
\begin{compactenum}
  \item We say that the $x$ is \emph{straight} if the path from $B_{(i-1,j)}$ to $B_{i+1,j}$ in $T_x$ contains all the vertices of $P_x$.
  \item Otherwise, we say that $x$ is \emph{bent}.
\end{compactenum}
\begin{figure}
  \begin{center}
    \begin{tabular}{c@{\hspace{1cm}}c}
      \includegraphics{figs/bx-paths-1.pdf} & \includegraphics{figs/bx-paths-3} \\
      (a) & (b)
    \end{tabular}
  \end{center}
  \caption{The vertex $x:=(i,j)$ is (a)~straight if the path from $B_{(i,j-1)}$ to $B_{(i+1,j)}$ uses all vertices of $P_x$ or (b)~bent otherwise.}
  \label{straight_or_bent}
\end{figure}
Note that, if $x$ is straight, then the path from $B_{i,j-1}$ to $B_{i,j+1}$ in $T_x$ also contains all the vertices of $P_x$.  Furthermore, if $x$ is bent, then, for each $\overline\i,\overline\j\in\{-1,1\}$, the path from $B_{i,j+\overline\j}$ to $B_{i+\overline\i, j}$ in $T_x$ contains all the vertices of $P_x$.

\begin{lem}
  For any $k\times k$ pseudogrid $H$, $\chilin(G)\in\Omega(k)$.
\end{lem}

@@

    \item Let $S\subset\{1,\ldots,C\}$ be the set of colours for which Stage~1 succeeds and let $\overline{S}:=\{1,\ldots,C\}\setminus S$ be the set of colours for which Stage~1 fails.  Let $R_1:=\bigcup_{c\in S}\{x_c,y_c\}$ be the set of vertices and edges chosen in $R_1$.

@@

    First observe that, for each $p\in R_1$, \cref{voronoi_mapping}(\ref{never_far}) implies that $|\vor^{-1}(p)|\le|B_{7\beta/4}(p)|\le (7\beta/2+1)^2$.  
    \todo[inline]{Fuck! We run into problems here with large colour sets.  The next line is not true, for example.   We get a $C$ in the denominator because the only degree bound we have on $R_1$ is $C(7\beta/2+1)^2$.}
    This implies that, for any $A_0\subseteq\overline{S}$,
    \[
        |N_I(A_0)| \ge \frac{a|A_0|}{(7\beta/2+1)^2} \enspace .    
    \]  
    This has two implications:
    \begin{compactenum}
      \item For each $p\in R_1$,  $\deg_{I}(x) \le C(7\beta/2+1)^2$.
      \item Since each colour $c\in A$ is frequent, $\deg_I(c)\ge \alpha/(7\beta/2+1)^2$.
    \end{compactenum}
    These two conditions imply that, for each $A_0\subseteq A$, $|N_I(A_0)|\ge \alpha/\zeta^2\beta^4$.   In particular, for any $\alpha$ and $\beta$ with $\alpha > 2\zeta^2\beta^4$,  $|N_I(A_0)|\ge 2|A_0|$.  Therefore, by \cref{hall_vees}, there exists a subgraph $\Lambda$ of $I$ in which each vertex of $A$ has degree $2$ and each vertex of $B$ has degree at most one.  Therefore, each $c\in A=\overline{S}$ has two $\Lambda$-neighbours $p$ and $q$ in $B$.  The edge $cp$ is present in $I$ because $G_k$ contains a vertex $x_c$ with $s(x_c)=p$ and $c\in \phi(x_c)$.  Similarly, the edge $cq$ is present in $I$ because $G_k$ contains a vertex $y_c$ with $s(y_c)=q$ and $c\in\phi(y_c)$.  In this way, the graph $\Lambda$ defines the pairs $\{x_c,y_c\}$ for each $c\in\overline{S}$.
  \end{enumerate}

  Summarizing our work thus far, we have identified a subset  $R:=\bigcup_{c=1}^C\{x_c,y_c\}$ of at most $2c$ vertices and edges of $G_k$ where $c\in \phi(x_c)\cap \phi(y_c)$ for each $c\in\{1,\ldots,C\}$. Ultimately, we want to use these to find a path in $H$ that contains at least two vertices of each colour.  We start by first establishing some properties of $R\subseteq V(G_k)$.

  Let $R_1:=z\in\bigcup_{c\in S}\{x_c,y_c\}$ and $R_2:=z\in\bigcup_{c\in \overline{S}}\{x_c,y_c\}$ be the subsets of $R$ obtained during Stages~1 and 2, respectively. The following claim follows immediately from the algorithm used to select pairs in Stage~1.
  \begin{clm}\label{r1_far}
      For each distinct pair $z,z'\in R_1$, $\dist_{G_k}(z,z')\ge \beta$.
  \end{clm}

  The next claim states that any element $z'\in R_2$ is quite far from every element of $R_1$ except possibly its nearest neighbour in $R_1$.

  \begin{clm}\label{far_from_home}
    For each $x\in R_1$ and $y\in R_2$ with $\nu(y)\neq x$, $\dist_{G_k}(x,y)\ge \beta /(2+\delta)$.
  \end{clm}

  \begin{proof}
    Since $\nu(y)\neq x$, there exists some other $z\in R_1\setminus\{y\}$ with $\nu(y)=\nu(z)$.  Since $x,z\in R_1$, \cref{r1_far} implies that $\dist_{G_k}(x,z)\ge \beta$.  Since $\nu(y)=z$, $\dist_{G_k}(y, z) \le (1+\delta)\gamma(y)\le (1+\delta)\dist_{G_k}(y, x)$.  By the triangle inequality
    \[ \beta \le \dist_{G_k}(x,z)
        \le \dist_{G_k}(x,y)+\dist_{G_k}(y,z)
        % \le \dist_{G_k}(x,y) + (1+\delta)\dist_{G_k}(x,y)
        \le (2+\delta)\dist_{G_k}(x,y) \enspace . \qedhere
    \]
  \end{proof}

  \begin{clm}
    Any subset of $R$ containing containing $4$ elements contains a pair $(x,y)$ such that $\dist_{G_k}(x,y)\ge \psi\beta$, for some $\psi>0$ that depends only on $\delta$.
  \end{clm}

>>>>

For a graph $G$ and a vertex $v\in V(G)$, $\deg_G(v):=|\{vw\in E(G)\}|$ denotes the degree of $v$ in $G$.  For a any $S\subseteq V(G)$, $\deg_G(S):=\sum_{v\in S}\deg_G(v)$ is the total degree of $S$.

For a vertex $v$ in a graph $G$, the \emph{open neighbourhood} of $v$ is $N_G(v):=\{w:vw\in E(G)\}$ and the \emph{closed neighbourhood} of $v$ is $N_G[v]:=\{v\}\cup N_G(v)$.  For any $S\subseteq V(G)$, $N_G(s):=\bigcup_{v\in S}N_G(v)$ and $N_G[s]:=\bigcup_{v\in S}N_G[v]$.  The subgraph $G[S]$ of $G$ \emph{induced} by $S$ is the graph with vertex set $V(G[S]):=S\cap V(G)$ and edge set $E(G[S]):=\{vw\in E(G):\{v,w\}\subseteq S\}$.

For a graph $G$ and a partition $\mathcal{P}$ of $V(G)$, the \emph{quotient graph} $G/\mathcal{P}$ is the graph with vertex set $\mathcal{P}$ and that contains an edge $vw$ if and only $G$ contains an edge with one endpoint in $v$ and one endpoint in $w$.  When $\mathcal{P}$ consists of disjoint subsets of $V(G)$ that do not necessarily cover $V(G)$ then we use the convention that $G/\mathcal{P}:=G/\mathcal{P'}$ where $\mathcal{P}'$ is the completion of $\mathcal{P}$ obtained by introducing each missing element of $V(G)$ as singleton set.  More precisely $\mathcal{P'}:=\mathcal{P}\cup\{\{v\}:v\in V(G)\setminus(\cup P)\}$.

A \emph{subdivision} of a graph $G$ is any graph $G'$ that can be obtained from $G$ by replacing each edge $vw$ of $G$ with a path $v,s_1,s_2,\ldots,s_\ell,w$ in which each of the (newly introduced) internal vertices $s_1,\ldots,s_\ell$ have degree exactly $2$.  Each vertex in $V(G')\setminus V(G)$ is called a \emph{subdivision vertex}.  A $\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has exactly $\ell$ internal vertices.  For any graph $G$, we use $G^{\onesub}$ to denote the $1$-subdivision of $G$.  Note that there is a natural bijection between the vertex set $V(G^{\onesub})$ and the vertex/edge set $V(G)\cup E(G)$.  A $\le\!\!\ell$-subdivision of $G$ is a subdivision of $G$ in which each edge of $G$ is replaced by a path that has at most $\ell$ internal vertices.

\todo[inline]{Surely this next lemma is well-known?}
We make use of the following generalization of Hall's Marriage Theorem \cite{hall:on}:
\begin{lem}\label{hall_vees}
  Let $G$ be a bipartite graph with vertex parts $A$ and $B$ and having the property that, for any $A_0\subseteq A$, $|N_G(W)|\ge 2|A_0|$.  Then $G$ contains a subgraph $\Lambda$ in which each vertex of $A$ has degree exactly two and each vertex of $B$ has degree at most one.
\end{lem}

@@

  Let $A_0\subseteq A$ and $B_0\subseteq B$ be the subsets of $A$ and $B$, respectively, that can be reached from $v_0$ by $\Lambda$-alternating paths.
  Observe that, for any $x\in B_0$,  there is exactly one edge $vx$ of $\Lambda$ incident on $x$ and the other endpoint $v$ of this edge is in $A_0$.  Let $a$ be the number of edges of $\Lambda$ having one endpoint $A_0$ and one endpoint in $B_0$, so that $a=\deg_\Lambda(B_0)=|B_0|$.  Let $b$ be the number of edges of $\Lambda$ having one endpoint in $A_0$ and one endpoint in $B\setminus B_0$, so that $\deg_\Lambda(A_0)=a+b = |B_0|+b$.    Since $\deg_{\Lambda}(v) \le 2$ for each $v\in A_0$ and $\deg_\Lambda(v_0)<2$, $\deg_\Lambda(A_0)<2|A_0|$.  This yields the desired contradiction, since $|N_\Lambda(A_0)| \le \deg_\Lambda(A_0) = |B_0|+b = \deg_\Lambda(A_0)< 2|A_0|$.greater than
\end{proof}

\begin{rem}
  The constant $2$ that appears twice in \cref{hall_vees} can be replaced with any positive integer $c$.
\end{rem}

@@

@@

\begin{proof}
  Let $P:=\{v_1,\ldots,v_t\}$ where $p_i^{\leftup}:=(x_i,y_i)$.  Let $X:=\bigcup_{i=1}^t \{x_i\bmod 7,(x_{i+1})\bmod 7\}$ and let $Y:=\bigcup_{i=1}^t\{y_i\bmod 7,(y_i+1)\bmod 7\}$.  Then $|X|,|Y|\le 2t\le 6$.  In particular, there exists some $x\in\{0,\ldots,6\}\setminus X$ and some $y\in\{0,\ldots,6\}\setminus Y$. For each $i,j\in\Z$, let $c_{i,j}:=(7i+x+3,7j+y+3)$ and observe that the ball $B_{3}(c_{i,j})$ does not have any point of $P$ on its boundary.  Furthermore, the set of balls $S:=\{B_{3}(c_{i,j}):i,j\in\Z\}$ is a partition of $V(G^{\onesub}_{k})$.  Therefore $S$ is a set of pairwise-disjoint balls, each of radius $3$, such that each vertex in $P$ is contained in the interior of some ball in $S$.
\end{proof}

\begin{lem}
  Let $P$ be a subset of $V(G^{\onesub}_k)$ such that no ball of radius $18$ contains more than $3$ vertices in $P$.  Then there exists a set $S$ of disjoint balls, each of radius $3$ and such that each point of $P$ is in the interior of some ball in $S$.
\end{lem}

\begin{proof}
  Define a graph $H$ with vertex set $V(H):=P$ and that contains an edge $pq$ if and only if $\|pq\|_\infty \le 12$. Observe that, if some connected subgraph of $H$ contains $4$ vertices, then these $4$ vertices are contained in a ball of radius $18$.  Therefore each component of $H$ contains at most $3$ vertices.

@@

\begin{lem}\label{one_path_enchilada}
  Let $k$ be a positive integer, let $I,L,Q$ be disjoint subsets of $V(G^{\onesub}_{k})\setminus\partial G^{\onesub}_k$ that have the following properties:
  \begin{compactenum}[(1)]\setcounter{enumi}{-1}
    \item Any ball of radius 18 contains at most $3$ elements of $I\cup L\cup Q$.
    \item Each vertex in $I\cup L$ has degree $4$ and each vertex in $Q$ has degree $2$.
    \item No vertex in $V(G^{\onesub}_{10})$ is adjacent to three vertices of $Q$.
    \item No vertex in $I\cup L$ is adjacent to two vertices of $Q$.
  \end{compactenum}
  Then $G^{\onesub}_k$ contains a path $P$ that
  \begin{compactenum}[(a)]
    \item contains all the vertices in $I\cup L\cup Q$;
    \item goes straight through each vertex in $I$; and
    \item turns at each vertex in $L$.
  \end{compactenum}
\end{lem}

\begin{proof}
  Start with the snakelike path $P_0$ that visits every $7$th row of $G_k$. Cover the elements of $I\cup L\cup Q$ with disjoint balls of side length $3$.  Each such ball is intersected by exactly one row of the snakelike path.  Use \cref{seven_by_seven} to locally deform $P_0$ to pick up the objects in each ball that it intersects.
\end{proof}

@@

  We now argue that $\vor'$ satisfies (\ref{three_in_ball}).  Let $v_1,\ldots,v_4\in Q$ and $1\le i_1<\cdots<i_4\le|P|$ be such that $\vor'(v_j)=p_{i_j}$ for each $j\in\{1,\ldots,4\}$.  Suppose, for the sake of contradiction, that $i_1,\ldots,i_4$ are all distinct and that $v_1,\ldots,v_4$ are contained in some ball $B_{\beta/12}(c)$.  By the triangle inequality $\|v_iv_j\|_\infty \le \|v_ic\|_\infty+\|cv_j\|_\infty \le \beta/6$ for each  $i,j\in\{1,\ldots,4\}$.
  Since $v_1\in Q$, $d(v)\le\beta$ so
  \[
    \|p_{i_1}v_1\|_\infty \le d(v_1) + 3\beta/4 \le \beta + 3\beta/4 \enspace .
  \]
  Since $\vor'(v_2)=p_{i_2}\neq p_{i_1}$,
  \[
    \|p_{i_2}v_2\|_\infty
     < \|p_{i_1}v_2\|_\infty - 3\beta/4
     \le \|p_{i_1}v_1\|_\infty + \beta/6 - 3\beta/4
     \le \beta + \beta/5 \enspace .
  \]
  Since $\vor'(v_3)=p_{i_3}\neq p_{i_2}$,
  \[
     \|p_{i_2}v_3\|_\infty
     < \|p_{i_2}v_3\|_\infty - 3\beta/4
     \le \|p_{i_2}v_2\|_\infty + \beta/6 - 3\beta/4
     < \beta + 2\beta/6 - 3\beta/4 \enspace .
  \]
  Finally, since $\vor'(v_4)=p_{i_4}\neq p_{i_3}$,
  \[
    \|p_{i_4}v_4\|_\infty
    < \|p_{i_3}v_4\|_\infty - 3\beta/4
    \le \|p_{i_3}v_3\|_\infty + \beta/5 - 3\beta/4
    < \beta + 3\beta/6 - 3\beta/2 = 0 \enspace ,
  \]
  which is clearly a contradiction.  This establishes that $\vor'$ satisfies Condition~(\ref{three_in_ball}).\todo{TODO: Optimize and cleanup. Finish proving (iii)}
\end{proof}

@@

\begin{proof}
  (This proof only uses the fact that $G_{a\times b}$ has maximum degree  $4$.)  Since $G$ contains a $G_{a\times b}$ minor then, by definition, $G$ contains a subgraph $H$ and a partition $\mathcal{B}:=(B_x:x\in V(G_{a\times b}))\}$ of $V(H)$ such that
  \begin{inparaenum}[(i)]
    \item $B_x$ is non-empty and $H[B_x]$ is connected for each $x\in V(G_{a\times b})$;
    \item $H/\mathcal{B}$ is isomorphic to $G_{a\times b}$ using the isomorphism given by $B_x\mapsto x$.
  \end{inparaenum}
  Let $(H,\mathcal{B})$ be chosen so that $|E(H)|$ is minimum, over all choices of $(H,\mathcal{B})$ that satisfy (i) and (ii).

  An edge of $vw$ of $H$ is a \emph{border edge} if $v\in B_x$ and $w\in B_y$ for distinct $x,y\in V(G_{a\times b})$.  For each $x\in V(G_{a\times b})$, let $\partial B_x$ denote the set of border edges incident on one vertex of $B_x$.  For each $x\in V(G_{a\times b})$, let $T_x:= H[B_x]\cup \partial B_x$.   Since $H$ is edge-minimal, $|\partial B_x|=\deg_{G_{a\times b}}(x)\le 4$, for each $x\in V(G_{a\times b})$ and $T_x$ is a tree with exactly $\deg_{G_{a\times b}}(x)$ leaves.\todo{Check the preceding claim.}  We define $\mathcal{P}:=(P_x:x\in V(G_{a\times b}))$ as follows:
\begin{compactenum}
  \item \label{one_special} If $T_x$ has exactly one node $v$ of degree at least three, then $P_x:=\{v\}$.
  \item \label{two_degree_three} If $T_x$ has two nodes $v$ and $w$ each of degree $3$ then $P_x$ contains all vertices on the path in $T_x$ from $v$ to $w$.
  \item \label{just_a_path} Otherwise $T_x$ is a path $v_0,\ldots,v_\ell$ for some $\ell\ge 2$ and $P_x:=v_1$.
\end{compactenum}
It is straightforward to verify that, with this assignment $\mathcal{P}:=(P_x:x\in V(G_{a\times b}))$, the quotient graph $H/\mathcal{P}$ is a subdivision of $G_{a\times b}$, so that $H$ is indeed an $a\times b$ pseudogrid.
\end{proof}

The second case (\ref{two_degree_three}) in the proof of \cref{pseudogrid_minor} occurs when $\deg_{G_{a\times b}}(x)=4$, so $x=(i,j)$ for some $i\in\{2,\ldots,a-1\}$ and $j\in\{2,\ldots,b-1\}$.  This case is especially important in our proof because it corresponds to a case in which some of the vertices of $P_x$ are shared by the $i$th column and the $j$th row of the pseudogrid. However, a path in the pseudogrid that uses vertices of $B_x$ may or may not contain all the vertices of $P_x$ depending on whether or not the corresponding path in $G_{a\times b}$ bends at $x$.  This motivates the following definitions (see \cref{straight_or_bent}):
\begin{compactenum}
  \item We say that the $x$ is \emph{straight} if the path from $B_{(i-1,j)}$ to $B_{i+1,j}$ in $T_x$ contains all the vertices of $P_x$.
  \item Otherwise, we say that $x$ is \emph{bent}.
\end{compactenum}
\begin{figure}
  \begin{center}
    \begin{tabular}{c@{\hspace{1cm}}c}
      \includegraphics{figs/bx-paths-1.pdf} & \includegraphics{figs/bx-paths-3} \\
      (a) & (b)
    \end{tabular}
  \end{center}
  \caption{The vertex $x:=(i,j)$ is (a)~straight if the path from $B_{(i,j-1)}$ to $B_{(i+1,j)}$ uses all vertices of $P_x$ or (b)~bent otherwise.}
  \label{straight_or_bent}
\end{figure}
Note that, if $x$ is straight, then the path from $B_{i,j-1}$ to $B_{i,j+1}$ in $T_x$ also contains all the vertices of $P_x$.  Furthermore, if $x$ is bent, then, for each $\overline\i,\overline\j\in\{-1,1\}$, the path from $B_{i,j+\overline\j}$ to $B_{i+\overline\i, j}$ in $T_x$ contains all the vertices of $P_x$.

\todo[inline]{I can't yet prove the full result, which is why the next lemma has some weird conditions.}

\begin{lem}
  %For any $k\times k$ pseudogrid $H$
  Let $H$ be a $k\times k$ pseudogrid and $\mathcal{P}:=(P_x:x\in V(G_{k}))$ be such that
  \begin{inparaenum}[(i)]
      \item for each $x\in V(G_k)$, $H[P_x]$ is a path of length at most $r$ and
      \item $H/\mathcal{P}$ is a $\le\!\! r$-subdivision of $G_k$.
  \end{inparaenum}
  Then $\chilin(G)\in\Omega(k/r)$.
\end{lem}

@@

    \item Let $S\subset\{1,\ldots,C\}$ be the set of colours for which Stage~1 succeeds and let $\overline{S}:=\{1,\ldots,C\}\setminus S$ be the set of colours for which Stage~1 fails.  Let $R_1:=\bigcup_{c\in S}\{x_c,y_c\}$ be the set of vertices and edges chosen in $R_1$.  By construction, $\min\{\|pq\|_\infty: \{p,q\}\in \binom{R_1}{2}\}\ge \beta$.

@@

    First observe that each vertex $x$ of $G^{\onesub}_k$ corresponds to a path $P_x$ in $H$ that contains at most $r$ vertices.\footnote{For some subdivision vertices of $G^{\onesub}_k$, $P_x$ may be empty.}  This implies that $|\Phi(x)|\le r$ for each $x\in V(G^{\onesub}_k)$.  Next observe that, for each $p\in R_1$, \cref{voronoi_mapping}(\ref{never_far}) implies that $|\vor^{-1}(p)|\le|B_{7\beta/4}(p)|\le (7\beta/2+1)^2$.  This implies that $\deg_I(p) \le r(7\beta/2+1)^2$ for each $p\in R_1$.

    This implies that, for any $A_0\subseteq\overline{S}$,
    \[
        |N_I(A_0)| \ge \frac{a|A_0|}{r(7\beta/2+1)^2} \ge 2|N_I(A_0)| \enspace .
    \]
    for appropriately chosen $\alpha$ and $\beta$ (with $\beta\in O(\sqrt{\alpha}))$.\todo{Specify precise condition.}  Therefore, by \cref{hall_vees}, there exists a subgraph $\Lambda$ of $I$ in which each vertex of $A$ has degree $2$ and each vertex of $B$ has degree at most one.  Therefore, each $c\in A=\overline{S}$ has two $\Lambda$-neighbours $p$ and $q$ in $B$.  The edge $cp$ is present in $I$ because $G_k$ contains a vertex $x_c$ with $s(x_c)=p$ and $c\in \Phi(x_c)$.  Similarly, the edge $cq$ is present in $I$ because $G_k$ contains a vertex $y_c$ with $s(y_c)=q$ and $c\in\Phi(y_c)$.  In this way, the graph $\Lambda$ defines the pairs $\{x_c,y_c\}$ for each $c\in\overline{S}$.
  \end{enumerate}

  Summarizing our work thus far, we have identified a subset  $R:=\bigcup_{c=1}^C\{x_c,y_c\}$ of at most $2c$ vertices and edges of $G_k$ where $c\in \Phi(x_c)\cap \Phi(y_c)$ for each $c\in\{1,\ldots,C\}$.
  Let $R_1:=z\in\bigcup_{c\in S}\{x_c,y_c\}$ and $R_2:=z\in\bigcup_{c\in \overline{S}}\{x_c,y_c\}$ be the subsets of $R$ obtained during Stages~1 and 2, respectively.

  \todo[inline]{There is a technical condition missing here.  To apply \cref{one_path_enchilada} we need the elements of $R$ to be distance at least $3$ from the boundary of $G^{\onesub}_k$.  When counting multiplicity, we shouldn't count occurrences of a colour that are close to the boundary.}

  Conditions~(\ref{never_far}) and (\ref{three_in_ball}) of \cref{voronoi_mapping} imply that each any ball of radius $\min\{\beta/8-1, \beta/12\}/2$ contains at most three elements of $R$.  Therefore, (for sufficiently large $\beta$) \cref{one_path_enchilada}, there exists a path $P$ in $G^{\onesub}_k$ that contains every vertex of $R$, that goes straight at every straight vertex of $R$ and that bends at every bent vertex of $R$. These latter two conditions guarantee that $V(P)$ contains $P_x$ for each $x\in R$. For each colour $c\in\{1,\ldots,C\}$, there are at least two vertices $x_c,y_c\in R$ such that $P_{x_c}$ and $P_{y_c}$ each contain some vertex of colour $c$.  Therefore the path $P$ contains two vertices of each colour.  Therefore $P$ has no center. Therefore $\varphi$ is not a linear centered colouring of $H$.  Since $\varphi$ is an arbitrary $\epsilon k$-colouring of $H$, this implies that $\chilin(H)\ge \epsilon k$.
\end{proof}
